---
title: 'Homework #2'
author: "Dougie Mundine" 
output:
  html_document: null
  toc: true
  toc_float: true
  pdf_document: default
---

```{r global_options, message= FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.height=5, fig.width=8, warning=FALSE, echo=FALSE,  tidy.opts=list(width.cutoff=60))
```


```{r include=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lubridate)
library(readr)
library(kableExtra)
```

wdm866 


------------------------------------------------------------------------

# **Problem #1 **


```{r}

profs = read.csv("profs.csv")

ggplot(profs) + geom_histogram(aes(x=eval), col="black", fill="green", binwidth=0.2) + labs(x="Course Evaluation Scores", title= "Overall Count of Course Evaluation Scores")

```

This graph shows the distribution of course evaluation scores of professors at UT. As you can see the rating of around 4.2-4.3 has the most of all of the course evaluation scores. This gives us an overview on how many professors are in the range of high evaluations scores and low evaluation scores. 



```{r}
ggplot(profs) + geom_boxplot(aes(x=eval), color="grey", fill="magenta") + facet_wrap('native') + labs(title="Distrubtion of Native Speakers Categorized by Yes or No and their Course Evaluation Range", x="Course Evaluations") + theme(axis.ticks.y=element_blank(), axis.text.y=element_blank())
```

This graph here represents the range of course evaluations of professors who are Native professors and are not. As we can see from the graphs the average course evaluation scores for native professors is higher of that of non- native professors. In the yes graph their are some outliers to the left whose course evaluation score is very low.



```{r}
ggplot(profs) + geom_histogram(aes(x=eval), color="grey", fill="red", bins = 25) + facet_wrap('gender') + labs(title="Distribution of Professors Course Evaluation Scores Categorized by Gender  ", x="Course Evaluations")
```

This data shows us two different graphs of male and female professors. It is the number of course evaluation scores for each gender. The graph to the right has the arguably better course evaluation scores than that of the left graph. The right graph is male professors and their scores, and the left graph is female professors and their scores. 


```{r}
ggplot(profs) + geom_point(aes(x=beauty, y=eval)) + labs(title= "Beauty Score and Evaluation Score", x="Beauty Score", y="Course Evaluation Score")
```

This plot shows us the beauty score of each professor and their evaluation score. The plot attempts to make a relationship between beauty and evaluation score. We can see that the most attractive professor's actually don't have the highest course evaluation scores. As shown their is one professor who has a negative beauty score but still has a high course evaluation score. 

# **Problem #2 **
```{r}
bikes = read.csv("bikeshare.csv")
bikes1 <- bikes %>%
  group_by(hr) %>%
  summarize(mean(total))
colnames(bikes1)[2]<-"total"

ggplot(bikes1) + geom_line(aes(x=hr, y=total))+ labs(title="Average Hourly Bike Rentals Across The Day", x="Hour", y="Average Bike Rentals")


```

This line graph shows the average bike rentals in each hour across a span of a day. As shown in the graph there are a couple of peaks in bike rentals across the day. The hour after 7.5 and the hour before 17.5 are both peaks where bike rentals are rented the most. What we can infer is that bike rentals are specifically higher in the morning around the 8 hour mark (8:00 am) and the 17 hour mark (5:00 pm). This makes a lot of sense because people may need to get to work and to get back home at around these times. 

```{r}
options(dplyr.summarise.inform = FALSE)


bikes2 <- bikes %>%
  
  group_by(hr, workingday) %>%
  summarize(avg_total=mean(total, na.rm = TRUE))
   bikes2<-bikes2 %>%
  mutate(workingday=ifelse(workingday==1, "yes", "no"))
ggplot(bikes2) + geom_line(aes(x=hr, y=avg_total))+ facet_grid('workingday')+labs(title="Average Bike Rentals Across Each Hour of the Day", x="Hour", y="Average Total Bike Rentals")
```

This line graph shows the average bike rentals across each hour of the day with the limitation of if the day is a working day (graph yes) or not a working day (graph no). As one can see on a working day the average bike rentals across the 7th hour and 17th hour are peaks. Which I believe we can now say that this is because of the 9-5's people are working on a workday. Now as we can see on a not working day the average bike rentals is kind of a trend with a rising slope at from the 8th hour to the 12th hour. After the 12th hour there is just a smooth trend indicating that people may just be riding the bikes around for fun. We learned that bike rentals on a working day is peaked at certain hours because of how a certain jobs work and on a not working day rentals are smooth because ultimately people may just be having fun. 

```{r}
bikes3<- bikes%>%
  filter(hr==9)%>%
  group_by(workingday, weathersit)%>%
  summarize(avg_total = mean(total, na.rm= TRUE))
    bikes3 <- bikes3%>%
      mutate(workingday=ifelse(workingday==1, "yes", "no"))
    ggplot(bikes3, aes(x=weathersit, y=avg_total)) + geom_bar(stat='identity')+ facet_grid('workingday')+labs(title="Average Bike Ridership for Each Weather Situation in the 9th Hour", x="Weather Situation ", y="Average Bike Ridership")
```

This bar graph shows the average bike ridership in the 9th hour of the day with some circumstances. The circumstances are certain weather conditions (1,2,3) and if the day is a working day (yes) or not a working day (no). The bar graph shows when in weather situation 1 both days (working day and not working day) have high bike ridership. In weather situation 2 it is similar. However, in weather situation 3 bike ridership is down by almost 50% for both days. We can conclude that in weather situation 1 and 2 aren't as serious as 3. We can also infer that because of the working day there are still going to be more ridership because people need to get to work. We have learned that ridership is going to reduce on whether or not the severeness of the weather situation as well as if the day is a working day or not. 

# **Problem #3 **

```{r}
metro <- read.csv("capmetro_Ut.csv")

boarding_data <- data.frame(
  hour = sample(0:23, 1000, replace = TRUE),
  day_of_week = sample(c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"), 1000, replace = TRUE),
  month = sample(c("September", "October", "November"), 1000, replace = TRUE),
  boardings = rpois(1000, lambda = 50) 
)
average <- metro %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarise(avg_boardings = mean(boarding), .groups = 'drop')
average$day_of_week <- factor(average$day_of_week, levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
average$month <- factor(average$month, levels = c("Sep", "Oct","Nov"))


ggplot(average, aes(x = hour_of_day, y = avg_boardings, color = month)) +
  geom_line() +
  facet_wrap(vars(day_of_week), scales = "free_y") +  
  scale_color_manual(values = c("Sep" = "orange",   
                   "Oct" = "blue",     
                   "Nov" = "red")) +
  labs(
    title = "Average Boardings by Hour of the Day, Faceted by Day of the Week",
    x = "Hour of the Day",
    y = "Average Boardings",
    color = "Month",
    
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

These plots show the average boardings by hour, day of week, and month (Sep, Oct, Nov). 
 The hour of peak boardings are similar across the days and months. We can verify this with evidence on the graph looking at the hours of the days. They look similar at  15-17.5 hour. There are some peaks on Saturday that aren't the same as the rest of the graphs. I think the average boardings on Mondays in September look lower compared to other days and months because in September there is labor day which class is cancelled and fewer people are going to be using metro at UT. I think for the low average boardings on Wednesday, Thursday, and Friday in November is because of Thanksgiving and no one is really using metro they are with their families. 

```{r}
ggplot(metro, aes(x = temperature, y = boarding, color = weekend)) +
  geom_point(alpha = 0.4) + 
  facet_wrap(~ hour_of_day, ncol = 4) +
  labs(
    title = "Scatter Plot of Boardings and Temperature",
    x = "Temperature (F)",
    y = "Number of Boardings",
    color = "Type of Day",
    
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12, face = "bold"), 
    legend.position = "bottom"
  )

``` 


These figures shows number of boardings related to temperature (). There is not a noticable change. The graph stays relatively flat.


# **Problem #4 **

```{r}
 bill <- read.csv("billboard.csv")
top <- bill %>%
  group_by(performer, song) %>%
  summarize(count = n(), .groups = 'drop') %>%
  arrange(desc(count)) %>%
  slice_head(n = 10)



kbl(top) %>% kable_paper(full_width = T, html_font = "Cambria") %>% kable_styling(bootstrap_options = c("striped","hover","condensed", "responsive"))
caption <- "The table shows the top 10 most popular songs on the Billboard Top 100 since 1958, based on the total number of weeks a song spent on Billiboard Top 100."
cat(caption)

```



```{r, message=FALSE, warning=FALSE}
billboard_data <- read_csv("billboard.csv",show_col_types = FALSE)
musical <- billboard_data %>%
  filter(year != 1958 & year != 2021) %>%  
  group_by(year) %>%
  summarize(unique_songs = n_distinct(song), .groups = 'drop')  
ggplot(musical, aes(x = year, y = unique_songs)) +
  geom_line(color = "red") +
  labs(
    title = "Musical Diversity of the Billboard Top 100 Over Time",
    x = "Year",
    y = "Number of Unique Songs",
    
  ) +
  theme_minimal()
  


```


This graph shows the number of unique songs that appeared on the Billboard Top 100 chart each year; however, not including 1958 and 2021. The interesting part about this graph is the rise in unique songs in 2020. I figured there would be a lot of unique songs in the 90s and not in the 2010s. I believe this because I think the songs now sound similar and the songs back then sound way different.



```{r echo=FALSE}
billboard_data <- read_csv("billboard.csv",show_col_types = FALSE)

ten_week_hits <- billboard_data %>%
  group_by(performer, song) %>%
  summarize(weeks_on_chart = n(), .groups = 'drop') %>%
  filter(weeks_on_chart >= 10)  
ten_week_hits_count <- ten_week_hits %>%
  group_by(performer) %>%
  summarize(ten_week_hits = n(), .groups = 'drop') %>%
  filter(ten_week_hits >= 30)  
ggplot(ten_week_hits_count, aes(x = reorder(performer, ten_week_hits), y = ten_week_hits)) +
  geom_bar(stat = "identity", fill = "orange") +
  coord_flip() +  
  labs(
    title = "Artists with the Most Ten-Week Hits",
    x = "Artist",
    y = "Number of Ten-Week Hits",

  ) +
  theme_minimal()



```

This bar plot shows the number of ten-week hits for artists who have achieved  30 ten-week hits or more on the Billboard Top 100 since 1958.




